\chapter{Introduction to Mean curvature flow}

We introduce the second fundamental form and mean curvature associated with an immersion of a hypersurface. The mean curvature flow is then the negative gradient flow of the volume functional on hypersurfaces. 
\section{Fundamentals of hypersurfaces}

Let $M^{n}$ be a smooth $ n $-dimensional manifold with a smooth immersion $ X : M^{n} \to \R^{n+1} $. If $ X $ is a diffeomorphism onto its image, we say $ X $ is an \textbf{embedding}  and its image $ \mathcal{M}^{n} = X(M^{n}) $ has the structure of a smooth $ n $-dimensional submanifold of $ \R^{n+1} $. We say that $ M^{n} $ is an \textbf{immersed hypersurface} and $ \mathcal{M} $ is an \textbf{embedded hypersurface} respectively. Throughout this thesis, we will denote the embedded manifold $ X(M) $ by script $ \mathcal{M} $ to differentiate between the domain and its image. Let $ (U, \{x^{i}\}) $ be a coordinate system on $ M^{m} $, in Euclidean coordinates the pushforward of tangent vectors will be 
    \[ dX( \partial_{i}) \define \frac{ \partial X}{ \partial x^{i}} = \partial_{i}X \]
where $ dX : TM^{n} \to T\R^{n+1} $ is the derivative of $ X $. Since $ dX $ is an injection for each point in $ M^{n} $, we can define an inner product on $ TM^{n} $ which in local coordinates is given by 
    \[ g( \partial_{i}, \partial_{j}) = \left< \partial_{i}X, \partial_{j}X \right> \]
where $ \left< \cdot, \cdot \right> $ denotes the standard inner product on Euclidean space. Further, we can define the Levi-Civita connection on $ M^{n} $ from the Levi-Civita connection on $ \R^{n+1} $. Let $ X_{p} \in T_{p}\Rn $ be a vector and $ Y: U \to T\Rn|_{U} $ be a local vector field in a neighborhood $ U $ containing $ p $. The Levi-Civita connection of $ Y $ with respect to $ X $ on  $ \R^{n+1} $ is given by 
\[ D_{X_{p}}Y =  X_{p}(Y^{i}) \partial_{i}\]
where $ Y = (Y^{1}, \ldots , Y^{n+1}) $ are the components of $ Y $ in the standard coordinates. Using the immersion condition, we define a connection on $ TM^{n} $ induced from $ D $.  Let $ x \in M^{n} $ and $ u \in T_{p}M^{n}, \tilde{v} \in TM^{n}|_{U}$ for some open set $ U $ containing $ x $. Define a connection $ \nabla $ by 
\begin{equation}
    dX(\nabla_{u}\tilde{v}) = D_{dX(u)}(\tilde{V}) \label{connection}
\end{equation}
where $ \tilde{V} $ is an extension of $ dX(\tilde{v}) $ to an open set of $ \Rn $ containing $ X(U) $.  
\begin{lemma}
    The connection defined by \cref{connection} is well-defined and is the unique Levi-Civita connection on $ (M^{n},g) $.
\end{lemma}
When $ X $ is an embedding, the restriction of the tangent bundle of $ T\Rn|_{\mathcal{M}} $ can be decomposed as the direct sum 
\[ T\mathcal{M}\oplus N\mathcal{M}  \]
where $ N\mathcal{M} $ is the \textbf{normal bundle}  which can be described as 
\[ N\mathcal{M} = \{(p,\nu) \in T\Rn|_{ \mathcal{M}} : \left< u, \nu \right> =0 \text{ for all } u \in T_{p} \mathcal{M}\}. \]
For dimension reasons, the normal bundle at each point is one-dimensional. We fix a choice of unit normal $ \nu_{p} $ for each $ p \in \mathcal{M} $. This leads to \textbf{tangential projection} $ \cdot^{T} : T\Rn \to T \mathcal{M}$  and \textbf{normal projection} $ \cdot^{\perp} : T\Rn \to N \mathcal{M} $ maps of vectors in $ T\Rn $ given by 
\[ u^{T} = u-\left< u,\nu \right> \nu,\quad \text{ and }\quad u^{\perp} = \left< u,v \right>\nu\]
respectively. We can define the Levi-Civita connection on an embedded hypersurface $ \mathcal{M} $ using the normal projection, 
\[ \nabla_{u}V  = (D_{u}V)^{T}\]
where $ u  $ is a vector and $ V $ is a local vector field. Notice that this is consistent with \cref{connection} since $ dX^{-1} $ is the tangential component. The next step is to calculate the Christoffel symbols of the connection $ \nabla $. For local coordinates $ (U, \{x^{i}\}) $ in $ M^{n} $, the Christoffel symbols $ \Gamma_{ij}^{k}: U \to \R $ are obtained by the formula
\[ \nabla_{ \partial_{i}X} \partial_{j}X = \partial_{i} (\partial_{j}X) + \Gamma_{ij}^{k} \partial_{k}X \]
which simplifies to 
\[ \Gamma_{ij}^{k} \partial_{k}X = \partial_{i}( \partial_{j}X) - (\partial_{i}( \partial_{j}X))^{T} = (\partial_{i}( \partial_{j}X))^{\perp}\]



\section{Mean Curvature Flow}

Now we define the Mean Curvature Flow (MCF) on hypersurfaces. 
\begin{defn}
    A one-parameter family of immersion $ X: M^{n} \times I \to \Rn $ is said to evolve by \textbf{Mean Curvature Flow} (MCF) if 
		\begin{equation}
            \frac{ \partial}{ \partial t}X(p,t) = \vec{H}(X(p,t))  = -H(X(p,t))\nu(X(p,t)) \quad \forall (p,t) \in M^{n} \times I.
        \end{equation}
\end{defn}

Notice that the mean curvature vector $ \vec{H} = -H \nu  $ is independent of the direction of normal $ \nu $. %It will be proven in \cref{embeddings} that if $ X(\cdot,0) $ is an embedding then $ \X(\cdot,t) $ remains an embedding for all $ t \in I $. This allows us to talk about the mean curvature flow of 
The following lemma demonstrates the similarity of mean curvature flow with the heat equation 

\begin{lemma}
    The mean curvature vector is equal to the Laplace-Beltrami operator of the hypersurface  
    \[ \vec{H} = -H \nu = \Delta_{\mathcal{M}}X . \]
\end{lemma}
\begin{proof}
    Notice that $ \partial_{i}\partial_{j}X = \Gamma_{ij}^{k}\partial_{k}X - h_{ij}\nu $. Contracting this, 
    \begin{align*}
        \Delta_{\mathcal{M}}X & = g^{ij}\nabla_{i}\nabla_{j}X \\
        & = g^{ij}(\partial_{i}\partial_{j}X - \Gamma_{ij}^{k}\partial_{k}X) \\
        & = - H \nu
    \end{align*}
    
\end{proof}

\subsection{Examples of the mean curvature flow}

It is difficult to solve the Mean curvature flow PDE on an arbitrary hypersurface. The limited number of examples come from ansatz or special cases,
\begin{enumerate}
    \item \textbf{Shrinking spheres}: Let $ \mathbb{S}^{n}(r) \subset \Rn $ be sphere of dimension $ n $ with radius $ r $. Since the mean curvature $ H = \frac{n}{r} $ is constant across the sphere, we make the ansatz that the hypersurface remains spherical under mean curvature flow. Let $ \mathcal{M}_{t} = \mathbb{S}^{n}(r(t)) $ be the solution, then the PDE is reduced to an ODE given by \begin{equation}
        \frac{d}{dt}r(t) = -\frac{n}{r(t)}
    \end{equation}
    whose solution is $ r(t) = \sqrt{r_{0}^{2}-2nt} $ with $ r(0)=r_{0} $. So the shrinking spheres $ \mathbb{S}^{n}(\sqrt{r_{0}^{2}-2nt})$ are a solution to the mean curvature flow for $ t \in [0,\frac{r_{0}^{2}}{2n}) $. 
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[x=0.5cm,y=0.5cm,z=0.3cm,>=stealth]
            \draw[->] (xyz cs:x=-5) -- (xyz cs:x=5) node[above] {$x$};
            \draw[->] (xyz cs:y=-5) -- (xyz cs:y=5) node[right] {$z$};
            \draw[->] (xyz cs:z=-5) -- (xyz cs:z=5) node[above] {$y$};
            \shade[ball color = gray!40, opacity = 0.5] (0,0) circle (1cm);
            \draw (0,0) circle (1cm);
            \draw (-1,0) arc (180:360:1 and 0.3);
            \draw[dashed] (1,0) arc (0:180:1 and 0.3);
            \fill[fill=black] (0,0) circle (1pt);
            \shade[ball color = gray!40, opacity = 0.4] (0,0) circle (2cm);
            \draw (0,0) circle (2cm);
            \draw (-2,0) arc (180:360:2 and 0.6);
            \draw[dashed] (2,0) arc (0:180:2 and 0.6);
            %\draw[dashed] (4,0) arc (60:120:8);
            %\draw (-4,0) arc (240:300:8);
              
            \shade[ball color = gray!40, opacity = 0.6] (0,0) circle (0.5cm);
            \draw (0,0) circle (0.5cm);
            \draw (-0.5,0) arc (180:360:0.5 and 0.15);
            \draw[dashed] (0.5,0) arc (0:180:0.5 and 0.15);
            \fill[fill=black] (0,0) circle (0.5pt);
            \draw [black, -stealth, thick] (0,4) -- (0,3);
            \draw [black, -stealth,thick] (2.8,2.8) -- (2.3,2.3);
            \draw [black, -stealth,thick] (4,0) -- (3,0);
            \draw [black, -stealth,thick] (0,-4) -- (0,-3);
            \draw [black, -stealth,thick] (-4,0) -- (-3,0);
            \draw [black, -stealth,thick] (-2.8,2.8) -- (-2.3,2.3);
            \draw [black, -stealth,thick] (2.8,-2.8) -- (2.3,-2.3);
            \draw [black, -stealth,thick] (-2.8,-2.8) -- (-2.3,-2.3);
        \end{tikzpicture}
        \caption{Shrinking spheres of dimension 2}
    \end{figure}
    \item \textbf{Evolution of Graphs}: Let $ f : \R^{n} \to \R $ be a smooth function. The graph of $ f $ in $ \Rn $, 
    \[ \mathcal{M} = \{(x,f(x)) \in \Rn :  x \in \R^{n} \} \]
    is a smooth hypersurface. The mean curvature vector at $ (x,f(x)) $ for the hypersurface can be calculated to be,
    \[ \sqrt{1+ |\nabla f|^{2}} \, \text{div} \left( \frac{\nabla f}{\sqrt{1+ |\nabla f|^{2}}} \right).\]
    Ecker and Huisken proved in \cite{ecker1989mean} that graphs evolving under the mean curvature flow remain graphs. So a family of graphs $ \mathcal{M}_{t} = \{(x,f_{t}(x)) :  x \in \Rn\} $ with the condition 
    \[ \frac{\partial}{ \partial t}f_{t}(x) = \sqrt{1+ |\nabla f_{t}|^{2}} \, \text{div} \left( \frac{\nabla f_{t}}{\sqrt{1+ |\nabla f_{t}|^{2}}} \right) \]
    is a solution of the mean curvature flow. 
    \item \textbf{Minimal surfaces}: Minimal surfaces are the critical points of the volume functional. A hypersurface $ \mathcal{M} $ is said to be a \textbf{minimal hypersurface} if it satisfies $ H(x) = 0 $ for all $ x \in \mathcal{M}  $. Hence, minimal hypersurfaces are stationary solutions of the mean curvature flow. 
    \item \textbf{Products of solutions with Euclidean space}
    : Suppose $ \mathcal{M}_{t}^{n} \subset \Rn $ is a solution of the mean curvature flow. It is easy to verify that the mean curvature vector of the product $ \mathcal{M}_{t}^{n} \times \R^{m} \subset \R^{n+1} \times\R^{m}  $ is given by 
    \[ \vec{H}(x,y) = (H(x) \nu(x), 0), \]
    which implies that the time-parametrized product $ \mathcal{N}_{t} = \mathcal{M}_{t} \times \Rn $ is a solution of the mean curvature flow as well.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[x=0.5cm,y=0.5cm,z=0.3cm,>=stealth]
            \coordinate (O) at (0,0,0);
              \coordinate (A) at (2,0,0);
              \coordinate (B) at (0,2,0);
              \coordinate (C) at (0,0,2);
              \draw[->] (xyz cs:x=-5) -- (xyz cs:x=5) node[above] {$x$};
              \draw[->] (xyz cs:y=-5) -- (xyz cs:y=5) node[right] {$z$};
              \draw[->] (xyz cs:z=-5) -- (xyz cs:z=5) node[above] {$y$};
            %\draw[-latex] (O) -- (A) node[below] {$y$};
            %\draw[-latex] (O) -- (B) node[above] {$z$};
            %\draw[-latex] (O) -- (C) node[below] {$x$};
            \node[cylinder, draw, shape aspect=0.3, 
              cylinder uses custom fill, cylinder end fill=gray!50, 
              minimum height=0.7cm,
              cylinder body fill=gray!25, opacity=0.5, 
            scale=5, rotate=90]{};
            \draw [black, -stealth] (1.25,3.55) -- (0.5,3.55);
            \draw [black, -stealth] (-1.25,3.55) -- (-0.5,3.55);
            %\draw [-stealth] 
        \end{tikzpicture}
        \caption{Cylinder $ S^{1} \times \R $}
    \end{figure}  
\end{enumerate}
\subsection{Mean curvature flow as the gradient of the area functional}

Let $ \mathcal{M}_{0} \subset \Rn $ be a smooth hypersurface and $ X: M^{n} \times (-\epsilon, \epsilon) \to \Rn $ be a variation with $ X(\cdot,0) = \mathcal{M}_{0} $. Considering area as a function of time over the variation, we get \begin{equation}
    \frac{d}{dt}\text{Area}( \mathcal{M}_{t}) = \int_{ \mathcal{M}_{t}} \left< \partial_{t}X, H \nu \right> 
\end{equation}
Using this, the gradient of the area functional is 
	\[ \nabla \text{Area} = H \nu \]
	so the most efficient way to reduce the volume is to choose the variation so that 
	\[ \partial_{t}X = -\nabla \text{Area} = -H \nu \]
	which is the mean curvature flow. In particular, we get the following equation for the evolution of area under mean curvature flow,
	\[ \frac{d}{dt} \text{Area}( \mathcal{M}_{t}) =  - \int_{\mathcal{M}_{t}}H^{2}\]
which is the steepest descent of area in the space of hypersurface up to speed-parametrization. 
\subsection{Short-time existence}
The following theorem about the uniqueness and existence of mean curvature flow is proved in \cite{andrews2022extrinsic}. 
\begin{thm}
    Short time existence: Let $ X_{0}: M^{n} \to \Rn $ be a smooth immersion of a compact manifold without boundary. There exists an $ \epsilon >0 $ and a smooth solution $ X : M^{n} \times [0, \epsilon) \to \Rn$ to MCF, with $ X( \cdot, 0) = X_{0} $. Moreover, the solution is unique.
\end{thm}

\section{Maximum principle}\label{max}
We can extend the maximum principle on Euclidean space to general Riemannian manifolds in the following fashion. 
\begin{lemma}[Scalar maximum principle]
    Let $ g(t) \in [0,T) $ be a $ 1 $-parameter family of Riemannian metrics on a closed manifold $ \mathcal{M}^{n} $ and $ \beta : \mathcal{M}^{n} \times [0,T) \to \R $ be a locally bounded function. Let $ u : \mathcal{M}^{n} \times [0,T) \to \R$ be a $ C^{2} $ function satisfying the following inequality 
    \[ \frac{ \partial}{ \partial t}u(x,t) \ge \Delta_{g(t)}u + \beta u\]
    If $ u(x,0) \ge 0 $ for all $ x \in \mathcal{M}^{n}$, then $ u(x,t) \ge 0 $ for all $ (x,t) \in \mathcal{M}^{n} \times [0,T) $.
\end{lemma}

\begin{proof}
    
\end{proof}

The above maximum principle can be further extended to tensors. This was done by Hamilton in \cite{hamilton1982three} in the context of Ricci flow. Let $ M = M_{ij}dx^{i}\otimes dx^{j} $ be a symmetric $ 2 $-tensor. We say $ M $ is non-negative if $ v^{T}Mv = M_{ij}v^{i}v^{j} \ge 0 $ for all vectors $ v $. Let $ N_{ij} = p(M_{ij},g_{ij}) $ be a tensor formed by contracting products of $ M_{ij} $ with itself using the metric. Also suppose that whenever $ v $ is a null-eigenvector of $ M_{ij} $ (i.e. $ M_{ij}v^{j} = 0 $), we have $ N_{ij}v^{i}v^{j} \ge 0 $. Then the following maximum principle holds

\begin{lemma}[Tensor maximum principle]
	Let $ g(t) \in [0,T)  $ be a $ 1 $-parameter family of Riemannian metrics on a closed manifold $ \mathcal{M}^{n} $. Let $ M_{ij} $ be a symmetric non-negative tensor evolving by the equation 
	
        \[ \frac{ \partial}{ \partial t}M_{ij} = \Delta M_{ij} + N_{ij}  \text{ for all } (x,t) \in  \mathcal{M}^{n} \times [0,T) \]
	where $ N_{ij} = p(M_{ij},g_{ij}) $ satisfies the null-eigenvector condition above. If $ M $ is non-negative at $ t=0 $, then it remains non-negative on $ [0,T) $.
\end{lemma}

\subsection{Applications to mean curvature flow}

The tensor maximum principle can be used to prove that convexity is preserved under mean curvature flow. 

\begin{lemma}
    If 
\end{lemma}
\section{Evolution equations}
To understand the properties of mean curvature flow it is essential to know the evolution of geometric quantities of the hypersurface. Let $X : M^n \times I \to \R^{n+1}$ be a smooth solution of mean curvature flow, so 
\[ \dou_t X(x,t) = \overrightarrow{H}(x,t)  = -H(x,t)\nu(x,t).\]
The induced metric on the hypersurface is given by $g = X^*(\delta)$ where $\delta$ is the flat metric on $\R^{n+1}$. This means that if $\{x^i\}$ are local coordinates on $M^n$, then the components of the induced metric are given by
\[ g_{ij} = \delta(X_*(\dou_i),X_*(\dou_j)) = \left\langle  \frac{\dou X}{\dou x^i}, \frac{\dou X}{\dou x^j} \right\rangle.  \]
\begin{lemma} Let $X: M^n \times I \to \R^{n+1}$ be a solution of mean curvature flow. Then the evolution equation of the metric, normal, second fundamental form and mean curvature is given by
\begin{equation}
    \dou_t g_{ij} = -2Hh_{ij}
\end{equation}
\begin{equation}
\dou_t \nu  = \nabla H
\end{equation}
\begin{equation}
\dou_t h_{ij} = \Delta h_{ij} - 2Hh_{il}g^{lm}h_{mj}+|A|^2h_{ij}
\end{equation}
\begin{equation}
    \dou_t H = \Delta H + |A|^2H \label{dtH}
\end{equation}
\end{lemma}
\begin{proof}
1. In local coordinates we have 
\begin{align*}
    \dou_tg_{ij} & = \partial_{t} \left\langle \partial_{i}X, \partial_{j}X \right\rangle \\
& = \left\langle \partial_{i}(\partial_{t}X), \partial_{j}X \right\rangle + \left\langle \partial_{i}X, \partial_{t}(\partial_{j}X) \right\rangle\\
& = \left\langle \partial_{t}(-H \nu), \partial_{j}X \right\rangle + \left\langle \partial_{i}X, \partial_{j}(-H \nu)\right\rangle \\
& = -H \left\langle \partial_{t} \nu, \partial_{j}X\right\rangle  - H \left\langle \partial_{i}X, \partial_{t}\nu\right\rangle \\
& = -2Hh_{ij}
\end{align*}
2. Since $\langle \nu, \nu \rangle = 1$, we have $2\langle \partial_{t}\nu, \nu \rangle = 0$, so the vector $\partial_{t}\nu$ is in the tangent plane of the hypersurface. We can write it as a linear combination of tangent vectors $\{\partial_{j}X\}$ to get
\begin{align*}
\partial_{t}\nu &= \left\langle \partial_{t}\nu, \partial_{i}X \right\rangle \partial_{j}Xg^{ij} = - \left\langle \nu, \partial_{i}\left( \partial_{t}X \right) \right\rangle\partial_{j}X g^{ij} \\
& = \left\langle \nu, \partial_{i}\left( H \nu \right) \right\rangle \partial_{j}Xg^{ij}\\
& = \partial_{i}H \partial_{j}X g^{ij} + H \left\langle \nu, \partial_{i}\nu\right\rangle \partial_{j}Xg^{ij}\\
& = \partial_{i}H\partial_{j}Xg^{ij} = \nabla H
\end{align*}
3. From the relations \[
\partial_{i}\partial_{j}X = \Gamma^{k}_{ij}\partial_{k}X	-h_{ij}\nu \quad \text{ and } \quad\partial_{j}\nu = h_{jl}g^{lm}\partial_{m}X\] we get 
\begin{align*}
\partial_{t} h_{ij} & = - \partial_{t} \left\langle\partial_{i}\partial_{j}X ,\nu\right\rangle\\
& = \left\langle\partial_{i}\partial_{j}(H \nu), \nu \right\rangle - \left\langle \partial_{i}\partial_{j}X, \partial_{l}H \partial_{m}Xg^{lm}\right\rangle \\
& = \partial_{i}\partial_{j}H + H \left\langle \partial_{i}\left( h_{jm}g^{ml} \partial_{l}X \right), \nu \right\rangle -\left< \Gamma_{ij}^{k}\partial_{k}X-h_{ij}\nu,\partial_{l}H \partial_{m}Xg^{lm} \right>\\
& = \partial_{i}\partial_{j}H - \Gamma_{ij}^{k}\partial_{k}H + Hh_{jm}g^{ml}\left< \Gamma_{il}^{p}\partial_{p}X- h_{il}\nu, \nu \right> \\
& = \nabla_{i}\nabla_{j}H- Hh_{il}g^{lm}h_{mj}.
\end{align*}
4. Utilizing the previous evolution equation with the product formula of derivatives, 
\begin{align*}
\dou_t H &= \dou_t(g^{ij}h_{ij}) = (\partial_{t}g^{ij})h_{ij}+ g^{ij}\partial_{t}h_{ij}\\
& = -g^{ik}(\partial_{t}g_{kl})g^{lj}h_{ij}+ g^{ij}(\Delta h_{ij} - 2Hh_{il}g^{lm}h_{mj}+|A|^2h_{ij})\\
& = -g^{ik}(-2Hh_{kl})g^{lj}h_{ij}+ \Delta(g^{ij}h_{ij}) - 2Hg^{ij}g^{lm}h_{il}h_{mj}+|A|^2H\\
& = 2H|A|^2 + \Delta H - 2H|A|^2 + |A|^2H\\
& =  \Delta H+ |A|^2H.
\end{align*}
\end{proof}
\begin{corollary}
If mean curvature is positive everywhere on the initial hypersurface, then it remains so throughout the flow.
\end{corollary}
\begin{proof}
We apply the maximum principle to the evolution equation of $H$.
\end{proof}
\begin{remark}
This property of mean curvature holds even when the hypersurface is embedded in an arbitrary Riemannian manifold with positive Ricci curvature. 
\end{remark}
\section{The Avoidance Principle}

Let $X_i: M^n_i \times  [0, T) \to \R^{n+1}, i = 1,2 $ be properly immersed solutions to mean curvature flow such that at least one of  $ M^n_1$ or $M^n_2$ is compact. 
If the hypersurfaces are disjoint initially, i.e. $X_1(M^{n_1},0)\cap X_2(M^n_2,0) = \phi$, then they remain so. Define the distance function $d: M^n_1 \times M^n_2 \times  [0,T) \to \R$ between the solutions by \[
d(x,y,t) = |X_2(y,t) - X_1(y,t)|
.\] 
as a function of time. From the assumptions of compactness and properness, $d_0 \define \inf_{(x,y) \in M^n_1 \times M^n_2} d(x,y,0) >0$
\begin{figure}[h]
\centering
\begin{tikzpicture}
    \draw [red] plot [smooth cycle] coordinates {(3,0)(0,3)(-3,0)(0,-3)};
    %\draw [red](0,0) circle (3cm);
    \fill [red!20] plot [smooth cycle] coordinates {(3,0)(0,3)(-3,0)(0,-3)};
    
    
    \draw [green] plot [smooth cycle] coordinates {(2.5,0)(0,2.5)(-2.5,0)(0,-2.5)};
    \fill [green!20] plot [smooth cycle] coordinates {(2.5,0)(0,2.5)(-2.5,0)(0,-2.5)};
    \draw [dashed,green] (2.5,0) arc (60:120:5cm);
    \draw [green] (-2.5,0) arc (240:300:5cm);
    \draw [thick,green][-stealth](2.75,0) -- (2,0);
    \draw [thick,red][-stealth](3,0) -- (2.5,0);
    \draw [dashed,red] (3,0) arc (60:120:6cm);
    \draw [red] (-3,0) arc (240:300:6cm); 
\end{tikzpicture}
\caption{Disjoint hypersurfaces flowing under MCF}
\end{figure}
\begin{thm}
If $X_1$ and $X_2$ are solutions to mean curvature flow on closed manifolds with $ d_{0}>0$, then $ d(x,y,t)>0$ for all $ x,y,t$. In particular, if $ X_1(x,0) \neq X_2(y,0)$ for all $x \in M^n_1  $ and $ y \in M^n_2$, then $ X_1(x,t) \neq X_2(y,t)$ for all $ x \in M^{n}_1$, $y \in  M^n_2$ and $ t \in  [0,T)$. 
\end{thm}
\begin{proof}
Assume on the contrary $ d(x,y,t)$ is not everywhere strictly greater that $ d_{0} $. Then there exists a $ t_{0} $ such that $ d(x_{0},y_{0},t_{0}) = d_{0} - \delta $
\end{proof}
\begin{remark}
From the proof, we can conclude that the distance between the hypersurfaces is a non-decreasing function. Another way to see this is that $ d_{t} $ satisfies a heat-type parabolic equation 
    \[ TO\, DO \]
 on which maximum principle is applicable. 
\end{remark}
\begin{proof} Fix $\epsilon > 0$ and suppose that $e ^{\epsilon (1+t)}d(x,y,t)$ is not strictly greater than $d_0$. That is there exists a time $ t_0 > 0$ such that $e^{\epsilon(1+ t)}d(\cdot, \cdot, t)$ reaches $ d_0$. That is,\[
e^{\epsilon(1+t)}d(\cdot,\cdot,t)> 0 \quad \text{ for } t < t_0
.\]  and \[
e^{\epsilon(1+t)}d(x_0,y_0,t_0) = d_0 \quad \text{ for some }(x_0,y_0) \in   M^n_1 \times  M^n_2		
.\] Then \[
\pt (e^{\epsilon(1+t)}d)|_{(x_0,y_0,z_0)} \le 0 
.\] Let $D$ denote the Euclidean directional derivative, by $ \nabla_i$ the covariant derivative induced on $M^n_i$ by $X_i$, and by $\nabla $ the covariant derivative on $ M^n_1 \times  M^n_2$ induced by $ \nabla_1$ and $ \nabla_2$, we find \[
\nabla^2
.\] 

\end{proof}
\begin{remark}
We can phrase the avoidance principle as disjointness is an immortal property and jointness is an ancient property.
\end{remark}


\section{Long time existence}

In this subsection, we will prove that the blow-up of the second fundamental form is the only obstruction to continuing the flow. The proof goes by contrapositive, relying on the Bernstein-type estimates. This technique is very similar to the one Hamilton used for Ricci flow \cite{hamilton1982three}. A solution of the mean curvature flow $ X : M^{n} \times [0,T) \to \Rn $ is said to be \textbf{maximal}  if given any other solution $ Y : M^{n} \times [0,S) \to \Rn $ which coincides with $ X $ for $ t \in [0,T) \cap [0,S) $ we have $ T \ge S $. Such a $ T $ is said to be the \textbf{maximal time} for $ X $. 

\begin{thm}\label{longtimeexistence}
    Let $ X: M^{n} \times [0, T) \to \Rn $ be a solution of the mean curvature flow with $ M^{n} $ compact. If $ X $ is maximal then $ T< \infty $ and 
    \[ \sup_{M^{n} \times [0,T)} |A| = \infty.\]
\end{thm}

Before jumping into the proof we need a general notation for complicated tensor expressions occurring in evolution equations. 
\begin{defn}
    Given any two tensors $ A $ and $ B $, we write $ A * B $ to denote any linear combination of tensors formed by contraction on $ A_{i_{1} \dots i_{k}}^{a_{1} \dots a_{p}}B_{j_{1} \dots j_{l}}^{b_{1} \dots b_{q}} $ with $ g $ or $ g^{-1} $. The iterated product $ A * B * C $ can be viewed as $ A * (B * C) $ which is associative and can be written without brackets. Also, denote the multifold product $ \underbrace{A * \dots * A}_{p-\text{times}}$ by $ A^{*p} $. The Gauss equation in this notation yields $ \text{Rm} = A * A $ which after differentiation gives $ \nabla \text{Rm} = A * \nabla A $. 
\end{defn}

The following lemma will be necessary to find out the time derivative of covariant derivatives and gives the commutator relation between them. 
\begin{lemma}
    Let $ S $ be a tensor with an evolution equation given by 
    \[ \partial_{t} S = \Delta S + T \]
    where $ T $ is another tensor of the same rank. Then the evolution equation of the covariant derivative is \begin{equation}
        \partial_{t}\nabla S = \Delta \nabla S + A * A  * \nabla S + A * \nabla A * S + \nabla T.
    \end{equation}
\end{lemma}
\begin{proof}
    Recall the time evolution of Christoffel symbol is given by     
    \begin{equation}
        \partial_{t}\Gamma_{ij}^{k} = \frac{1}{2}g^{kl}(\nabla_{i}\partial_{t}g_{jl}+ \nabla_{j}\partial_{t}g_{il}-\nabla_{l}\partial_{t}g_{ij}).
    \end{equation}
    Substituting $ \partial_{t}g_{ij} = -2Hh_{ij} $ we get 
    \[ \partial_{t}\Gamma_{ij}^{k} = -g^{kl}(\nabla_{i}(Hg_{jl})+ \nabla_{j}(Hg_{il})-\nabla_{l}(h_{ij})) = A * \nabla A. \]
    Consider the commutator \begin{align*}
        \partial_{t}\nabla S & = \nabla \partial_{t}S + \partial_{t}\Gamma *S \\
        & = \nabla(\Delta S+T) + A * \nabla A * S \\
        & = \Delta \nabla S + \nabla \Rm * S + \Rm * \nabla S  + \nabla T + A * \nabla A * S \\
        & = \Delta \nabla S + A * A  * \nabla S + A * \nabla A * S + \nabla T
    \end{align*}
    where we have used the Ricci identity $ [\nabla, \Delta]S = \nabla \Rm * S + \Rm * \nabla A $.
\end{proof}



\begin{lemma}
    The evolution equation of the higher gradient of the second fundamental form is given by 
    \begin{equation}
        \partial_{t}\nabla^{m}A = \Delta \nabla^{m}A + \sum_{i+j+k =m}^{}\nabla^{i}A* \nabla^{j}A*\nabla^{k}A 
    \end{equation}
    where $ m \in \N \cup \{0\} $. Further, the norm of the gradient satisfies 
    \begin{equation}
        \partial_{t}|\nabla^{m}A|^{2} = \Delta |\nabla^{m}A|^{2}- 2 |\nabla^{m+1}A|^{2} +\sum_{i+j+k =m}^{}\nabla^{i}A* \nabla^{j}A*\nabla^{k}A* \nabla^{m}A \label{normnablaA}
    \end{equation}
    where $ m \in \N \cup \{0\} $.
    
\end{lemma}
\begin{proof}
    We induct on $ m $. For base case $ m = 0 $, the second fundamental form evolution equation is 
    \begin{align*}
        \partial_{t} A & = \Delta A -2H A^{2} + |A|^{2}A \\
        & = \Delta A + A*A*A. 
    \end{align*}
    Now suppose the equation holds for $ m $, then for $m +1 $ we have \begin{align*}
        \partial_{t}\nabla^{m+1}A & = \nabla \partial_{t} (\nabla^{k}A) + (\partial_{t}\Gamma)* \nabla^{m}A \\
        & = \nabla \left(\Delta \nabla^{m}A + \sum_{i+j+k=m}^{}\nabla^{i}A * \nabla^{j}A * \nabla^{k}A\right) + A * \nabla A * \nabla^{m}A \\
        & = \Delta \nabla^{m+1}A + \nabla \Rm * \nabla^{m}A + \Rm* \nabla^{m+1}A + \sum_{i+j+k =m+1}^{}\nabla^{i}A* \nabla^{j}A*\nabla^{k}A \\
        & = \Delta \nabla^{m+1}A + \sum_{i+j+k =m+1}^{}\nabla^{i}A* \nabla^{j}A*\nabla^{k}A
    \end{align*}
    using Ricci identity and Gauss equation $ \Rm = A*A $. For the norm, we get \begin{align*}
        \partial_{t}|\nabla^{m}A|^{2} = 2\left<\partial_{t}\nabla^{m}A, \nabla^{m}A  \right> + A*A* \nabla^{m}A * \nabla^{m}A
    \end{align*}
    where the second term comes from time derivative $ \partial_{t}g^{ij} = -2Hh^{ij} = A*A $. This simplifies to \begin{align*}
        \partial_{t}|\nabla^{m}A|^{2} & = 2\left< \Delta \nabla^{m}A + \sum_{i+j+k=m}^{}\nabla^{i}A * \nabla^{j}A* \nabla^{k}A, \nabla^{m}A  \right> + A*A*\nabla^{m}A * \nabla^{m}A \\
        & = 2\left< \Delta \nabla^{m}A , \nabla^{m}A \right> + \sum_{i+j+k=m}^{}\nabla^{i}A * \nabla^{j}A* \nabla^{k}A \nabla^{m}A \\
        & = \Delta|\nabla^{m}A|^{2} - 2|\nabla^{m+1}A|^{2} + \sum_{i+j+k=m}^{}\nabla^{i}A * \nabla^{j}A* \nabla^{k}A* \nabla^{m}A
    \end{align*}
\end{proof}
We can consider the maximum principle on the previous lemma. This gives control of derivates of the second fundamental form based on the bound of the second fundamental form. 

\begin{lemma}
    Let $ X: M^{n} \times [0, T) \to \Rn $ be a solution of the mean curvature flow with $ M^{n} $ compact. Suppose that $ T < \infty $ and $ C_{0} = \sup_{M^{n} \times [0,T)}|A| < \infty $. Then for each $ m \in \N $ there exists a constant $ C_{m} < \infty $ depending only on initial manifold such that \begin{equation}
        \sup_{M^{n} \times [0,T)} |\nabla^{m}A| \le C_{m}.
    \end{equation}
\end{lemma}

\begin{proof}
    From \cref{normnablaA}, we can estimate \begin{align*}
        \partial_{t}|\nabla^{m}A|^{2} \le \Delta|\nabla^{m}A|^{2} - 2|\nabla^{m+1}A|^{2} + C_{n,m}\sum_{i+j+k=m}^{}|\nabla^{i}A|  |\nabla^{j}A|| \nabla^{k}A|| \nabla^{m}A|
    \end{align*}
    where $ C_{n,m}< \infty $ only depends on $ n $ and $ m $. We now proceed by induction on $ m $. Suppose that for $ l \in \{1, \dots, m-1\} $ we have 
    \[ \sup_{\mathcal{M} \times [0,T)} |\nabla^{l}A|^{2} \le C_{l}\]
    for some constants $ C_{l} < \infty$. From the previous relation, \begin{align}
        \partial_{t}|\nabla^{m}A|^{2} & \le \Delta |\nabla^{m}A|^{2} - 2 |\nabla^{m+1}A|^{2} + C_{n,m} \left( |A|^{2} |\nabla^{m}A|^{2} + \sum_{\substack{i+j+k=m \\
        i,j,k \le m-1}}^{}|\nabla^{i}A||\nabla^{j}A||\nabla^{k}A||\nabla^{m}A|  \right) \nonumber \\
        & \le \Delta|\nabla^{m}A|^{2}+ C_{n,m}C_{0}^{2}|\nabla^{m}A|^{2} + c_{m}|\nabla^{m}A|  \nonumber \\
        & \le \Delta|\nabla^{m}A|^{2} + 2C_{n,m}C_{0}^{2}|\nabla^{m}A|^{2} + \frac{c_{m}}{4C_{0}^{2}C_{n,m}} \label{nablam}
    \end{align}
    using the induction hypothesis. To absorb the $ |\nabla^{m}A|^{2} $ term on the right-hand side consider the inequality for $ m-1 $, \begin{align}
        \partial_{t}|\nabla^{m-1}A|^{2} &\le \Delta |\nabla^{m-1}A|^{2} - 2|\nabla^{m}A|^{2} + C_{n,m-1} \sum_{i+j+k=m-1} |\nabla^{i}A||\nabla^{j}A| |\nabla^{k}A| |\nabla^{m-1}A| \nonumber\\
        & \le \Delta|\nabla^{m-1}A|^{2} - 2|\nabla^{m}A|^{2} + c_{m-1} \label{nablam1}.
    \end{align}
    Multiplying \cref{nablam1} by $ C_{n,m}C_{0}^{2} $ and adding to \cref{nablam}, \begin{align*}
        \partial_{t}\left( |\nabla^{m}A|^{2}+ C_{n,m}C_{0}^{2}|\nabla^{m-1}A|^{2} \right) & \le \Delta\left( |\nabla^{m}A|^{2}+ C_{n,m}C_{0}^{2}|\nabla^{m-1}A|^{2} \right) + \frac{c_{m}}{4C_{0}^{2}C_{n,m}} \\
        & \qquad + c_{m-1}C_{n,m}C_{0}^{2}.
    \end{align*}
    Now by comparison principle, we get a bound 
    \[ \sup_{\mathcal{M}^{n} \times [0,T)}\left( |\nabla^{m}A|^{2}+ C_{n,m}C_{0}^{2}|\nabla^{m-1}A|^{2} \right) \le \overline{C_{m}} \]
    from this, we can deduce that 
    \[ \sup_{\mathcal{M} \times [0,T)} |\nabla^{m}A|^{2} \le C_{m}\]
    for some constant $ C_{m} $ depending only on $ n,m, C_{0}, \dots, C_{m-1} $ and the initial hypersurface. This completes the induction.
\end{proof}
We can improve the higher-order covariant derivative bound to include a time factor as well. The following lemma will imply a rapid decrease in the norm of higher covariant derivatives of the second fundamental form with respect to time. 

\begin{lemma}
    Let $ X : M^{n} \times [0,r^{2}] \to \Rn  $ be a solution of the mean curvature flow. Suppose there exists a constant $ C_{0} < \infty$ such that 
    \[ |A|^{2} \le C_{0}r^{-2} \]
    on $ \mathcal{M}^{n} \times [0,r^{2}] $. Then for each $ m \in \N $, there exists a constant $ C_{m} $ depending only on $ n, \mathcal{M}_{0} $ and $C_{0} $ such that 
    \[ t^{m}|\nabla^{m}A|^{2} \le C_{m}r^{-2} \]
    on $ \mathcal{M}^{n} \times [0,r^{2}] $. 
\end{lemma}
\begin{proof}
    First, we will demonstrate how to obtain the bound for $ m=1 $ and then extend the method for general $ m $ using induction. We know that 
    \begin{align*}
        \partial_{t}|\nabla A|^{2} & = \Delta|\nabla A|^{2}-2|\nabla^{2} A|^{2} + \sum_{i+j+k=1}^{}\nabla^{i}A* \nabla^{j}A *\nabla^{k}A * \nabla^{m}A \\
        & = \Delta|\nabla A|^{2}-2|\nabla^{2} A|^{2} + C_{n,1} \sum_{i+j+k=1}^{}|\nabla^{i}A||\nabla^{j}A| |\nabla^{k}A||\nabla A|  \\
        & \le \Delta|\nabla A|^{2} -2 |\nabla^{2} A|^{2}+3C_{n,1}|A|^{2}||\nabla A|^{2} \\
        & \le \Delta|\nabla A|^{2} -2 |\nabla^{2} A|^{2}+3C_{n,1}C_{0}r^{-2}|\nabla A|^{2}.
    \end{align*}
    To obtain a better bound we want to utilize the good term $ -2|\nabla^{2}A|^{2} $ and one way to do this is to define another term with a $ t $-factor and $ |A|^{2} $,
    \[ F = t|\nabla A|^{2} + \beta|A|^{2}\]
    which is bounded at $ t=0 $ by $ \beta C_{0}r^{-2} $. Notice that \begin{align*}
        \partial_{t}F & \le |\nabla A|^{2}+ t\left(\Delta|\nabla A|^{2}-2|\nabla^{2} A|^{2} +3C_{n,1}C_{0}r^{-2}|\nabla A|^{2} \right) + \beta\left( \Delta |A|^{2}- 2|\nabla A|^{2} + 2|A|^{4} \right) \\
        & \le \Delta F -2|\nabla^{2}A|^{2}+  \left(1 +3tC_{n,1}C_{0}r^{-2}-2\beta  \right) |\nabla A|^{2} + 2\beta C_{0}^{2}r^{-4} \\
        & \le \Delta F + \left( 1+ 3C_{n,1}C_{0}^{2}-2 \beta \right)| \nabla A|^{2} + 2 \beta C_{0}^{2}r^{-4}.
    \end{align*}
    Choose  $ \beta > (1+3C_{n,1}C_{0}^{2}) /2 $, so that the coefficient of $ |\nabla A|^{2} $ is negative. Comparison theorem then gives 
    \[ \sup_{ x \in \mathcal{M} }F(x,t) \le \beta C_{0}r^{-2}+ 2 \beta C_{0}^{2}r^{-4}t \le Cr^{-2} \]
    for some constant $ C_{1} >0 $. Hence, 
    \[ t|\nabla A|^{2} \le C_{1}r^{-2} \]
    on $ \mathcal{M} \times [0,r^{2}] $ which establishes the inequality for $ m=1 $. 

    Now assume the inequality holds for $ 1, \dots,m-1 $. Then, 
    \begin{align*}
        \partial_{t}|\nabla^{m} A|^{2} &\le \Delta |\nabla^{m}A|^{2}- 2|\nabla^{m+1}A|^{2} + C_{n,m} \sum_{i+j+k= m}^{} |\nabla^{i}A||\nabla^{j}A||\nabla^{k}A||\nabla^{m}A|  \\
        & \le \Delta |\nabla^{m}A|^{2}- 2|\nabla^{m+1}A|^{2} + C_{n,m}\left( 3|A|^{2}|\nabla^{m}A|^{2} + \sum_{\substack{i+j+k=m\\i,j,k \le m-1}}^{}|\nabla^{i}A||\nabla^{j}A| |\nabla^{k}A||\nabla^{m} A|  \right) \\
        & \le \Delta|\nabla^{m}A|^{2}- 2|\nabla^{m+1}A|^{2}+ 3C_{n,m}C_{0}r^{-2}|\nabla^{m}A|^{2} + C_{n,m}L_{m}r^{-3}t^{-\frac{m}{2}}|\nabla^{m}A| \\
        & \le \Delta|\nabla^{m}A|^{2}- 2|\nabla^{m+1}A|^{2} + + C_{n,m}r^{-2}\left(6C_{0}|\nabla^{m}A|^{2} + \frac{L_{m}^{2}r^{-2}t^{-m}}{4C_{0}}  \right) \\
        & \le \Delta|\nabla^{m}A|^{2}- 2|\nabla^{m+1}A|^{2} + r^{-2}(T_{m}|\nabla^{m}A|^{2}+ S_{m}r^{-2}t^{-m})
    \end{align*}
    where $ L_{m} = \sum_{\substack{i+j+k=m\\i,j,k \le m-1}}\sqrt{C_{i}C_{j}C_{k}}  $, $ T_{m} = 6C_{n,m}C_{0} $ and $ S_{m} = (C_{n,m}L_{m}^{2}) / 4C_{0} $.

    Now like in the $ m=1 $ case, we define 
    \[ F = \sum_{l=0}^{m}a_{l}t^{l}|\nabla^{l} A|^{2} \]
    where $ a_{l} $ are positive constants to be determined later. The time derivative of $ F $ satisfies \begin{align*}
        \partial_{t}F & \le \sum_{l=0}^{m}a_{l}t^{l}\left(\Delta|\nabla^{l}A|^{2}- 2|\nabla^{l+1}A|^{2} + r^{-2}(T_{l}|\nabla^{l}A|^{2}+ S_{l}r^{-2}t^{-l}) \right) + \sum_{l=0}^{m-1}(l+1)a_{l+1}t^{l}|\nabla^{l+1}A|^{2} \\
        & \le \Delta F + \sum_{l=0}^{m-1}((l+1)a_{l+1}-2a_{l}+)t^{l}|\nabla^{l}A|^{2} -2|\nabla^{m+1}A|^{2} + \sum_{l=0}^{m}a_{l}r^{-4}S_{l} 
    \end{align*}


\end{proof}

\begin{remark}
    The parabolic nature of the mean curvature flow shows up here as the dimension of time is double the spatial dimension coming from $ |\nabla^{m}A| $.
\end{remark}
\begin{proof}[ of \cref{longtimeexistence}]
    Assume on the contrary that $ \sup_{M^{n} \times [0, T)}|A|^{2} \le C $. We aim to show that the manifold $X(\cdot, t)= \mathcal{M}_{t} $ converges to a smooth limit $ \mathcal{M}_{T} $ as $ t \to T $ which will give a contradiction by applying the short time existence of the mean curvature flow at $ t  = T $. \begin{align*}
        |X(x,t)-X(x,0)| \le  \bigg|\int_{0}^{t} H(x,s) ds \bigg| \le T \sup |H| \le T \sqrt{nC_{0}}
    \end{align*}
    %hence $ X(t, \cdot) \to X(T, \cdot) $ converges to a continuous limit. 
\end{proof}










\section{Huisken's theorem}

Huisken's theorem states that compact, uniformly convex hypersurface converges to a round sphere under Mean curvature flow in finite time. 
\begin{thm}
Let $ X: M^{n} \times [0, T) \to \Rn  $, $ n \ge 2 $ be a maximal solution of MCF such that $ M^{n} $ is compact and $X_{0} = X(\cdot, 0)$ is convex embedding. Then $ X_{t} = X(\cdot,t) $ is a convex embedding for all $ t>0 $ and $ X_{t} $ converges to a point $ p \in \Rn $ as $ t \to T $. Further the rescaled embeddings $ \tilde{X}_{t} : M^{n} \to \Rn $ defined by 
            \[ \tilde{X}_{t}(x) \define \frac{X_{t}(x)-p}{\sqrt{2n(T-t)}}\]
converge uniformly in the smooth topology to a smooth embedding whose image coincides with the unit sphere $ S^{n} $.
\end{thm}
\subsection{Pinching estimate}

The quantity $ \frac{|A|^{2} - \frac{H^{2}}{n}}{H^{2}} = \frac{1}{n} \sum_{i<j}^{} \left( \frac{\kappa_{i}}{H}- \frac{\kappa_{j}}{H} \right)^{2}  $ is scaling invariant and measures the roundness of the hypersurface. If we compute the evolution equation we get \begin{equation}
EQUATION
\end{equation} 
but the maximum principle is not directly applicable because of the positive last term. So we do not directly get a pointwise $ L^{\infty} $ bound we do the next best thing possible which is $ L^{p} $ bounds. After obtaining this there is a sophisticated iteration argument developed by Stampacchia which allows us to produce an $ L^{\infty} $ estimate. See  \href{https://youtu.be/a577KPiOoxw?t=1790}{here}
\begin{thm}
There exists constants $ \delta $ and $ C_{0} < \infty $ depending only on $ \mathcal{M}_{0} $ such that 
\[ |A|^{2} - \frac{H^{2}}{n} \le C_{0}H^{2-\delta} \]
for $ t \in (0, T] $.

\end{thm}
Let $ f_{\sigma} = \frac{|A|^{2}- \frac{H^{2}}{n}}{H^{2-\sigma}} = \left(\frac{|A|^{2} - \frac{H^{2}}{n}}{H^{2}}\right)H^{\sigma}$, then the evolution equation of $ f_{\sigma} $ is 
\begin{lemma}
The evolution of $ f_{\sigma} $ is given by \begin{align*}
    \frac{ \partial}{ \partial t}f_{\sigma} & = 
\end{align*}
\end{lemma}
\begin{comment}
    (Huisken's remark on this from the lecture series: the bad term can only be controlled using integral estimates \href{https://www.mfo.de/about-the-institute/staff/prof-dr-gerhard-huisken/lectures/mean-curvature-flow/lecture-3}{58 min mark} )

(Sinestrari intuition about Stampacchia iteration : \href{https://youtu.be/XvUGowx9WNo?t=2119}{yt link})
\end{comment}
\begin{comment}
    
\subsection{Width-pinching}

Let $ \Omega  \subset \Rn  $ be a compact, convex body with boundary $  \mathcal{M}  = \partial \Omega$. %such that $ \mathcal{M} $ is locally uniformly convex. 
%This allows us to define the inverse of the Gauss map on the hypersurface $ \mathcal{M} $. 
Define the support function $ \sigma : S^{n} \to \R $ as follows 
\[ \sigma(z) \define \sup_{x \in \Omega}\left< x,z \right> .\]
Using this we define the width function of $ \Omega $ by 
\[ w(z) = \sigma(z)+ \sigma(-z) \qquad \text{ for all } z \in S^{n} \]
which as the name suggests measures the width of $ \Omega  $ in the direction $ z$. Let $ w_{+} $ and $ w_{-} $ denote the maximum and minimum width. It is easy to see that maximum width is the diameter of $ \Omega $, so 
\[ w_{+} = \sup_{x,y \in \Omega}||x-y||. \]
If $ \mathcal{M}$ is a smooth uniformly convex hypersurface, then we can write the support function in terms of Gauss map $ G: \partial \Omega \to S^{n} $ as 
\[ \sigma(z) = \left< G^{-1}(z), z \right>, \] so 
\[ \sigma(G(z)) = \left< z,G(z) \right>.\]

It is a lemma by Andrews that the pinching of principal curvatures implies the pinching of the widths, which will be later important to prove the convergence of the hypersurface into a point. 
\begin{lemma}
Let $ \Omega \subset \Rn $, $ n \ge 2 $ be a uniformly convex body with compact, smooth boundary $ \mathcal{M} = \partial \Omega $. Suppose the principal curvatures of $ \mathcal{M} $ are pinched by a constant $ C < \infty $, i.e. $ \kappa_{n}(x) \le C \kappa_{1}(x) $ for all $ x \in \mathcal{M}  $. Then the widths of $ \mathcal{M} $ are pinched by the same constant, so 
\[ w_{+} \le C w_{-} \]
\end{lemma}
\begin{proof}
We will first prove it for $ n=2 $, which presents the main idea. Let $ e  \in S^{2}$ be unit vector in $ \R^{3} $ and (TO DO LATER) 
\end{proof}

We know the solution of MCF of spheres, so it will be useful to relate the widths of the hypersurface to the spheres inscribed and circumscribed about it. Let $$ \rho_{+} \define \inf\{r : \Omega \subset B_{r}(x) \text{ for some }x \in \R^{n}\}$$ denote the circumradius and 
\[ \rho_{-} \define \sup\{B_{r}(X) \subset \Omega \text{ for some }x \in \R^{n}\} \]
denote the inradius radius. There is also a relation between inscribed and circumscribed radius to widths of the hypersurface as given by the following lemma
\begin{lemma}
On any bounded convex body $ \Omega \subset \Rn $, 
\[ \rho_{+} \le \sqrt{n+1}w_{+} \quad \text{ and } \quad \rho_{-} \ge \frac{1}{n+2}w_{-}.\]
\end{lemma}
\begin{remark}
These are not the best bounds however they are sufficient for our purpose.
\end{remark}

Combining the two lemmas, we get 
\begin{equation}
\rho_{+} \le \frac{n+2}{\sqrt{n+1}}C\rho_{-} \label{pinching}
\end{equation}

Let $ \{\mathcal{M}_{t}\}_{t \in [0,T)}$ be family of hypersurfaces flowing under MCF with maximal time $ T $ and the initial pinching constant given by
\[ C = \sup_{ \mathcal{M}_{0}} \frac{\kappa_{n}}{\kappa_{1}} \]

If we can show that $ \rho_{-} \to 0 $ as $ t \to T $, then inequality (\ref{pinching}) would imply $ \rho_{+} \to 0 $. This proves the convergence of hypersurfaces to a point. 
\begin{thm}
The inradius $ \rho_{t} $ of $ \Omega_{t} $ tends to zero as $ t \to T $. Thus, $ \mathcal{M}_{t} $ converges to some point $ p \in \Rn $ as $ t \to T $.
\end{thm}
\end{comment}

\section{Monotonicity Formula}
Let $ X: M^{n} \times I \to \R^{n+1} $ be a one-parameter family of immersions flowing by mean curvature. Let $ \{x^{i}\} $ be local coordinates around a point $ p \in M $. Then the metric and second fundamental form is given by 
\[ g_{ij} = \left< \frac{\dou X}{\dou x^{i}}, \frac{\dou X}{\dou x^{j}} \right> , \qquad h_{ij} = \left<\eta, \frac{\partial^{2}X}{\dou x^{i}\dou x^{j}} \right>. \]

If we scale the solution by a factor of $ \lambda  $, defined by $ \tilde{X}(x,t) = \lambda X(x,t) $ we get the following metric and second fundamental form 

\[ \tilde{g_{ij}} = \left< \frac{\dou \tilde{X}}{\dou x^{i}}, \frac{\dou \tilde{X}}{\dou x^{j}} \right>  = \lambda^{2}g_{ij} ,\qquad \tilde{h_{ij}} = \left<\eta, \frac{\partial^{2}\tilde{X}}{\dou x^{i}\dou x^{j}} \right>  = \lambda h_{ij}\]
so the scaled mean curvature is given by $ \tilde{H} = \tilde{g}^{{ij}}\tilde{h}_{ij} = \frac{1}{\lambda}H $. This implies the scaled solutions satisfy the evolution equation

\[ \frac{\dou \tilde{X}}{\dou t} = \lambda \frac{\dou X}{\dou t} = -\lambda^{2}\tilde{H}\eta \]
or
\[ \frac{\dou \tilde{X}}{\dou (\lambda^{2}t)} = - \tilde{H}\eta \]
Hence if scale time by $ \lambda^{2} $, then $ \tilde{X} $ is also a solution of the mean curvature flow.

Mean curvature flow is invariant under parabolic scaling, i.e.  if $ X : M^{n} \times I \to \R^{n+1} $ is solution, then so is $ X_{\lambda}(x,t) = \lambda X(x,\lambda^{2}t) $. We construct a weighted area functional which is invariant under \textit{parabolic} scaling along any solution to mean curvature flow which will be monotonous.

Let $ \rho(x,t) $ be the backward heat kernel at $ (X_{0},t_{0}) $, i.e.,   \[\rho(x,t) = \frac{1}{(4 \pi (t_{0}-t))^{ \frac{n}{2}} }\cdot \exp\left( - \frac{|X(x,t)-X_{0}|^{2}}{4(t_{0}-t)} \right), \qquad t<t_{0} \]
\begin{thm}
If $ M_{t} $ is a solution of mean curvature flow for $ t< t_{0} $, then we have the formula

    \[ \frac{d}{dt} \int_{M_{t}}\rho(x,t)d \mu_{t}  = - \int_{M_{t}}\rho(x,t)\left( H- \frac{\left< X(x,t)-X_{0}, \eta \right>}{2 (t_{0}-t)} \right)^{2}d \mu_{t}\] 

\end{thm}
\begin{proof}
To simplify the formula assume that $ \left(X_{0},t_{0}\right) = (0,0) $. We know that $ \frac{d}{dt} \mu_{t} = -H^{2} \mu_{t} $, so differentiating $ \rho $ with respect to time we get,
\begin{align}
    \frac{d}{dt} \int_{M_{t}}\rho(x,t)d \mu_{t}  &= \int_{M_{t}}\rho(x,t)(-H^{2})d \mu_{t} + \int_{M_{t}} \frac{\dou}{\dou t}\rho(x,t)d \mu_{t}\nonumber \\
    & = -\int_{M_{t}}\rho(x,t)H^{2}d \mu_{t} +\int_{M_{t}}\left( \frac{\left< X(x,t), H(x,t)\eta \right>}{2(-t)}\rho(x,t) \right)d\mu_{t}\nonumber \\
    &  \qquad \qquad  + \int_{M_{t}} \left( \frac{n}{2(4 \pi)(-t)}(4 \pi)\rho(x,t) -\frac{|X(x,t)|^{2}}{4(-t)^{2}}\rho(x,t)\right)d \mu_{t}\nonumber \\ 
    & = \int_{M_{t}}\rho\left(   \frac{n}{2(-t)} +\frac{\left< X,H \eta\right>}{2(-t)}- \frac{|X|^{2}}{4(-t)^{2}} -H^{2}\right)d \mu_{t}\label{rho0}
    %& = -\int_{M_{t}}\rho \left| H \eta + \frac{X}{2(-t)}\right|^{2}d \mu_{t} + \int_{M_{t}}\rho \frac{ \left< X,H \eta \right>}{2(-t)} d \mu_{t} + \int_{M_{t}} \frac{n\rho}{2(-t)}d \mu_{t} \label
\end{align}

Now $ \Delta X = - H \eta $, using this relation for the second term and divergence theorem we get 
\begin{align}
\int_{M_{t}}\rho \left< X,H \eta \right> d \mu_{t} & =  -\int_{M_{t}} \rho\left< X,\Delta X \right> d \mu_{t} \nonumber \\ 
& = - \sum_{k=1}^{n+1}\int_{M_{t}} \rho X_{k} \Delta X_{k} d \mu_{t}\nonumber \\
& = \sum_{k=1}^{n+1} \int_{M_{t}} \left< \nabla(\rho X_{k}), \nabla X_{k} \right>d \mu_{t}\nonumber \\ 
& = \sum_{k=1}^{n+1} \int_{M_{t}} \left( \left< \nabla\rho, \nabla X_{k} \right>X_{k} + \rho \left< \nabla X_{k}, \nabla X_{k} \right> \right)d \mu_{t} \label{rho1}
% & = \int_{M_{t}} \left< \nabla(\rho X), \nabla X \right>d \mu_{t}\\ 
% & = \int_{M_{t}} \left( \rho\left< \nabla X, \nabla X  \right> + \left< \nabla \rho, \nabla X \right> X \right)
\end{align}
Let $ (U,\{x^{i}\}) $ be some local coordinates on the hypersurface. In these coordinates we can write $ \nabla \rho = g^{ij} \partial_{i} \rho \partial_{j} $, so $  \left< \nabla \rho , \nabla X_{k}\right> = \nabla \rho (X_{k}) = g^{ij} (\partial_{i}\rho) (\partial_{j}X_{k}) $ which implies 
\begin{align}
\sum_{k=1}^{n+1} \left< \nabla \rho , \nabla X_{k}  \right> X_{k} & = \sum_{k=1}^{n+1}  g^{ij} (\partial_{i}\rho) (\partial_{j}X_{k}) X_{k}\nonumber \\ 
& = g^{ij}( \dou_i \rho) \left< X, \dou_j X \right>\nonumber \\
& = g^{ij} \rho \left( \frac{-\left< X, \partial_{i}X \right>}{2(-t)} \right) \left< X, \dou_j X \right>\nonumber \\
& = - \frac{\rho}{2(-t)}|X^T|^2 \label{rho2}
\end{align}
and 
\begin{equation}
\sum_{k=1}^{n+1} \rho  \left< \nabla X_{k}, \nabla X_{k} \right> = \sum_{k=1}^{n+1} \rho g^{ij} (\partial_{i} X_{k} )(\partial_{j}X_{k}) = \rho g^{ij} \left<  \partial_{i}X, \partial_{j}X \right> = \rho g^{ij}g_{ij} = n \rho \label{rho3}
\end{equation}
Substituting \cref{rho2} and \cref{rho3} into \cref{rho1} and multiplying by $ \frac{1}{2(-t)} $, we get 
\begin{equation*}
\int_{M_{t}}\rho  \frac{\left< X,H \eta \right>}{2(-t)} d \mu_{t}  = \int_{M_{t}} \rho \left( \frac{n}{2(-t)} - \frac{1}{4(-t)^2}|X^T|^2  \right) d \mu_{t} 
\end{equation*}
or 
\begin{equation}
\int_{M_{t}}\frac{n\rho}{2(-t)} d \mu_{t} = \int_{M_{t}} \rho \left(  \frac{\left< X,H \eta \right>}{2(-t)} + \frac{1}{4(-t)^2}|X^T|^2 \right)d \mu_{t}\label{rho4}
\end{equation}
where $ X^{T} $ denotes the tangential part of the vector $ X $. Substituting \cref{rho4} into \cref{rho0}
\begin{align*}
\frac{d}{dt} \int_{M_{t}}\rho(x,t)d \mu_{t}  &=  \int_{M_{t}}\rho\left(\frac{\left< X,H \eta\right>}{(-t)}- \frac{|X|^{2}}{4(-t)^{2}} -H^{2}+ \frac{1}{4(-t)^2}|X^T|^2\right)d \mu_{t}\\ 
& = -\int_{M_{t}} \rho \left( H - \frac{\left< X, \eta \right>}{2(-t)} \right)^{2} d \mu_{t}.
\end{align*}
\end{proof}

\subsection{Rescaled Monotonicity formula}\label{Type1singularity}

From section (?) we know that the curvature blows up at the maximal time $ T $  and satisfies the inequality

\[ \max_{p \in M}|A(p,t)| \ge \frac{1}{\sqrt{2(T-t)}} \]

\begin{defn}
Let $ T $ be the maximal time of existence of a mean curvature flow. If there exists a constant $ C >1 $ such that 

    \[ \max_{p \in M}|A(p,t)| \le \frac{C}{\sqrt{2(T-t)}} \]
we say the flow is developing at time $ T $ a \textit{type I singularity}.
\end{defn}

Conversely, if such a constant does not exist, that is 

\[ \limsup_{t \to T} \max_{p \in M}|A(p,t)|\sqrt{T-t} = \infty \]

we say that we have a \textit{type II singularity}. We will restrict ourselves to type I singularity for the rest of this section. 


\section{Surfaces of positive mean curvature}

From the maximum principle, we know that if the mean curvature of the initial hypersurface $M_0$ is positive then it will stay positive on $M_t$. 
For self-similar solutions, we know that the limiting hypersurface will satisfy the equation $H = \langle x, \nu \rangle$. 
We prove that the sphere is the only compact hypersurface of positive mean curvature moving under self-similarity 
\begin{thm}
If $M^n$, $n\ge 2$, is compact with non-negative mean curvature $H$ and satisfies the equation $H = - \langle X, \nu \rangle$, then $M^n$ is a sphere of radius $\sqrt{n}$. 
\end{thm}
\begin{proof}
Suppose the hypersurface satisfies $H = -\langle X, \nu \rangle$. Let $e_1,\ldots, e_n$ be an orthonormal frame on $M^n$, then 
\begin{align}
    \nabla_{i}H & = - \left< D_{e_{i}}X, \nu \right> - \left< X, \nabla_{e_{i}}\nu \right> \nonumber \\
    & = -\left< e_{i}, \nu \right> - \left< X, \left< \nabla_{e_{i}}\nu,e_{l} \right>e_{l} \right>\nonumber \\ 
    & = \left< X,e_{l} \right>h_{il}
\end{align}
\[\nabla_i \nabla_j H = h_{ij} - Hh_{il}h_{lj}+\langle x, e_l \rangle \nabla_l h_{ij}\]
\end{proof}
